Eliminating prior-bias from sparse-projection tomographic reconstructions

### 2020-07-20 paper/tci.tex
minor edit
edited figures,supplemetary material

 When the number of views is significantly reduced (this paper), the
 sparsity prior alone is not sufficient.  In such cases additional
 information specific to the object being scanned (i.e., object prior)
 is useful in further improving the reconstruction.  While this has
 been done in the context of dynamic CT-scans (as discussed in
 Section~\ref{sec:related}), in this paper we use a set $S$ of
 previous scans of the same object. This is relevant in a longitudinal
 context: the acquisition of sequential CT scans of the same subject
 in order to track time-evolving changes within the subject's
 interior. As seen in Fig.~\ref{fig:story} (last column, 30 views) our
 method uses the prior scans \emph{uniformly} by creating an
 eigenspace representation (Section~\ref{sec:method_uniform_prior}).

 However, with the use of such an object prior, a new challenge
 emerges. The prior set $S$ may potentially overwhelm the necessary
 details; when several prior scans are available, finding the right
 prior is an issue. We therefore seek an algorithm that estimates the
 location and magnitude of new changes in the (unknown) test. As we
 show in Section~\ref{sec:method_spatially_varying_prior}, this
 eventually prevents the prior from adversely affecting the
 reconstruction of new regions in the test. We refer to this method as
 a \textit{spatially-varying} prior-based reconstruction routine, that
 still uses \textit{all} the previous scans in the set $S$, without
 using simply the prior scan, or endeavoring to find the ``right'' prior.

@@

In this section, we survey the existing literature on tomographic reconstruction from undersampled measurements using object priors. In most cases, the object prior consists of an object reconstructed to high quality from a large number of projection angles. One of the earliest pieces of work in this area is the well-cited PICCS method~\cite{PICCS}, which enforces a robust-norm based similarity between the object to be reconstructed (called the `test' henceforth) and the object prior, in addition to a sparsity prior on the test. However in such a technique, there is no inherent mechanism to prevent the prior from overwhelming the current reconstruction, and choice of relative weights between the object and sparsity priors is difficult. A closely related method called PIRPLE was proposed in~\cite{pirple}, with additional steps to register the object prior and the test and a somewhat more flexible combination of data fidelity terms, sparsity prior and object prior. However, this method too has similar limitations as PICCS. In some pieces of work, the object prior is projected in the same angles as the current set of test measurements, and a set of projection differences is computed. These projection differences are essentially tomographic projections of the difference between the unknown test and the known object prior. Hence, a variety of reconstruction algorithms can be used to estimate such a difference image which can then be added to the object prior to yield the final estimate of the test. Such methods have been proposed in~\cite{Pourmorteza2015,Lee2012}. However, the reconstruction of the difference image will inherently contain artifacts due to a sparse set of projections, and these artifacts will appear in the final reconstruction as there is no mechanism to mitigate them (unlike the method that we propose). In some recent techniques such as~\cite{Marjolein2016}, additional knowledge regarding regions of most likely change between the test and object prior is assumed, and lower weights are assigned to the prior in those regions. However, this is not done in a data driven manner and relies on additional expert domain knowledge which may not be available or which may be infeasible to acquire. There exists a fairly large-sized body of literature for reconstruction of structurally changing objects. In such cases, it may be infeasible to gather a large number of projections at every time instant. Moreover, there is additional temporal redundancy available which can reduce the required sample complexity for good reconstruction, and at the same time, tracking structural changes across time is of paramount importance. Many interesting approaches for this have been proposed in~\cite{daniil2015,koen2020, Van2015,HaoGao,Van2014}. These approaches variously use 4D sparsifying transforms for exploiting spatio-temporal redundancy~\cite{daniil2015}, robust principal components analysis assuming that the object consists of a large static part (background) and a sparse moving part (foreground)~\cite{HaoGao}, optical flow computations between projections (at different time instances) of the dynamic nonrigid object~\cite{koen2020} with additional key-frame and residual modelling in~\cite{Guangming2019}, spline-based models for tracking the regions of change in an SIRT framework~\cite{Van2014}, or knowledge of fluid models (piecewise constant shapes of attenuation curves in the dynamic region) in an SIRT framework~\cite{HaoGao}. However all these techniques ~\cite{daniil2015,koen2020,Van2015,HaoGao,Van2014} essentially reconstruct all the different stages of the object simultaneously, and use the extra redundancy across time simultaneously. However, in the scenario of a longitudinal study (where projection measurements are acquired at instants that are several weeks apart) which we consider in this paper, such approaches are not feasible. A longitudinal study will necessarily require reconstruction results after acquisition of each set of measurements. 
In this paper, we present a technique that makes use of multiple object priors, corresponding to high quality reconstructions of similar objects across time. We use these priors to improve the reconstruction of the test from largely undersampled measurements. In particular, we define the regions of change between the test and the eigenspace spanned by the previous high quality object priors in a purely data driven manner. Very importantly, our technique clearly differentiates between genuine structural changes and ``changes'' that appear due to undersampling artefacts. Since we use a statistical model with multiple object priors, we avoid the problem of selecting an appropriate single object prior unlike~\cite{PICCS,Pourmorteza2015,Lee2012,pirple,Marjolein2016} and also make use of the additional information available in the multiple priors.

@@

 \begin{figure}[h]
\centering
	\includegraphics[width=0.5\textwidth]{../images/prior_cmb.png}
        \caption{Overview of our work. The choice of the number of
          measurement views and the type of reconstruction is driven
          by the goal in the application under consideration.  (a) When
          our goal is relatively simple, such as tracking the location
          of new changes while simultaneously reducing sub-sampling
          artefacts, we propose acquiring measurements from an
          extremely small number of views (`few-view' imaging) and
          using uniform prior based reconstruction. (b) When our goal
          becomes more ambitious, such as observing details of the new
          changes while simultaneously reducing sub-sampling
          artefacts, we propose acquiring measurements from a
          slightly higher number of views (`moderate-view' imaging)
          and using spatially-varying prior-based reconstruction. In
          either case, the number of views is lower than what is
          conventionally used.}
 \label{fig:prior_overview}
\end{figure}

In summary, the key contributions are
\begin{itemize}
\item We create new 3D biological datasets (Section~\ref{sec:datasets}) and
  present results on real cone-beam projections. Our datasets and code
  will be made available to the community.
  \item As seen in Fig.~\ref{fig:story}, we demonstrate the use of
    uniform priors when very few views are used. More results appear
    in Section~\ref{sec:results_uniform_prior} and in the
    supplementary material.
  \item We discuss the scenarios where a uniform object prior is
    sufficient for the application at hand, and the scenarios for
    which it will fail . This leads to the idea of a weights map
    designed to depict the location and strength of new changes at
    every voxel. A novel algorithm is presented to build this map from
    sub-sampled measurements of the test and a set of high quality
    templates. Once the weights map is built, it is used for accurate
    reconstruction of those changes in the test that are absent in all
    of the templates. Results appear in
    Section~\ref{sec:results_spatially_varying_prior}.
  \item We show the efficacy of our results in a real-life medical
    longitudinal studies with data obtained in a clinical setting from
    a live teaching and
    research hospital 
\end{itemize}

@@

 \textbf{Okra:} This dataset is that of an Okra specimen consisting of its five
scans (Figure~\ref{fig:object-prior_test_okra}). The measurements
consisted of real circular cone-beam projections from 450 views, each of size
336$\times$156. The corresponding size of the reconstructed volume is
338$\times$338$\times$123. Prior to the first scan, two holes were
drilled on the surface of the specimen. This was followed by four
scans, each after introducing one new cut. The specimen was kept in
the same position throughout the acquisitions. In cases where such an
alignment is not present, all the template volumes must be pre-aligned
before computing the eigenspace. The test must be registered to the
object-prior after its preliminary pilot reconstruction. The ground
truth consists of volumes reconstructed using the Feldkamp-Davis-Kress
(FDK) algorithm~\cite{FDK} from the full set of 450 view
projections. The test volume was reconstructed from a partial set of
45 projections, i.e., $10\%$ of the projection views from which ground
truth was reconstructed.\\
 %--------------------------------okra data-------------------------------
\begin{figure}[!h]
    \begin{subfigure}[b]{0.18\linewidth}
        \includegraphics[width=\textwidth]{../images/okra/templateCropped_1.png}
\captionsetup{labelformat=empty}       
 \caption{}
    \end{subfigure}
    \begin{subfigure}[b]{0.18\linewidth}
        \includegraphics[width=\textwidth]{../images/okra/templateCropped_2.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.18\linewidth}
        \includegraphics[width=\textwidth]{../images/okra/templateCropped_3.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.18\linewidth}
        \includegraphics[width=\textwidth]{../images/okra/templateCropped_4.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.176\linewidth}
        \fcolorbox{yellow}{yellow}{\includegraphics[width=\textwidth]{../images/okra/testCropped.png}}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
     \caption{Okra 3D dataset: One slice each from the previously scanned objects (the
       first four from the left), and one from the test volume
       (extreme right). In the regions marked in red and green, while
       all slices have deformities, the test
       has none.}
\label{fig:object-prior_test_okra}
\end{figure}
%----------------------------------------------------

\textbf{Potato:} This dataset consisted of four scans of the humble potato (Figure~\ref{fig:object-prior_test_potato_A}). Measurements from each scan consisted of real circular cone-beam projections from 900 views, each of size 150$\times$150. The corresponding size of the reconstructed volume is 150$\times$150$\times$100. While the first scan was taken of the undistorted potato, subsequent scans were taken of the same specimen, each time after drilling a new hole halfway into the potato.  Projections were obtained using circular cone beam geometry.  %Further, this alignment can be refined following every update of the reconstructed test volume.
The ground truth consists of FDK reconstructions from the full set of acquired measurements from 900 projection views. The test volume was reconstructed using measurements from 45 projection views, i.e., $5\%$ of the projection views from which ground truth was reconstructed.
%---------------------------------------
\begin{figure}[!h]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/template_1.png}
\captionsetup{labelformat=empty}       
 \caption{}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/template_2.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/template_3.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.235\linewidth}
        \fcolorbox{blue}{blue}{\includegraphics[width=\textwidth]{../images/potato/testIm_green.png}}
\captionsetup{labelformat=empty}
\caption{}
\label{fig:potato3D_test}
     \end{subfigure}
      \caption{Potato 3D dataset: One slice (slice-A) each from the previously scanned objects (the first three from left) and a slice from the test 
        volume (extreme right). Notice the appearance of the fourth
        hole in the test slice. }
\label{fig:object-prior_test_potato_A}
\end{figure}

\subsection{Simulated measurements from real-life reconstructed volumes}
\textbf{Sprouts:} This 3D dataset, which was also obtained at ANU, consists of six reconstructed
volumes corresponding to six scans of an in-vivo sprout specimen
imaged at its various stages of growth
(Figure~\ref{fig:object-prior_test_sprouts}). Projections were generated
from the given volume of size 130$\times$130$\times$130. The ground
truth consists of FDK reconstructed volumes from a set of 1800 view
projections. The test volume was reconstructed from partial set of 45
projections, i.e., $2.5\%$ of the projection views from which ground
truth was reconstructed.\\
%-------------------------sprouts data

\begin{figure}[!h]
    \begin{subfigure}[b]{0.3\linewidth}
        \includegraphics[width=\textwidth]{../images/sprouts/template_1.png}
\captionsetup{labelformat=empty}
        \caption{}
    \end{subfigure}
\quad
    \begin{subfigure}[b]{0.3\linewidth}
        \includegraphics[width=\textwidth]{../images/sprouts/template_2.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
\quad
    \begin{subfigure}[b]{0.3\linewidth}
        \includegraphics[width=\textwidth]{../images/sprouts/template_3.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}\\
    \begin{subfigure}[b]{0.3\linewidth}
        \includegraphics[width=\textwidth]{../images/sprouts/template_4.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
\quad
    \begin{subfigure}[b]{0.3\linewidth}
        \includegraphics[width=\textwidth]{../images/sprouts/template_5.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
\quad
    \begin{subfigure}[b]{0.29\linewidth}
        \fcolorbox{green}{green}{\includegraphics[width=\textwidth]{../images/sprouts/testIm_red.png}}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
      \caption{Sprouts 3D dataset: One slice each from the previously scanned objects (the first five from left) and a slice from the test (extreme
        right).}
\label{fig:object-prior_test_sprouts}
\addtolength{\textfloatsep}{-0.8cm}
\end{figure}

\textbf{Radio-frequency ablation:} We have used this dataset (Figure~\ref{fig:RFA2_test_object-prior} to
illustrate an application where \textit{both} uniform and spatially-varying priors are
useful. The data~\footnote{Source: Tata Memorial
  Centre~\cite{tmh}, Parel, Mumbai.  This is the national
  comprehensive centre for the prevention, treatment, education and
  research in cancer, and is recognized as one of the leading cancer
  centres in India.} consists of simulated parallel beam measurements of 2D slices
of the organ (liver) at different stages of the scan. The details of this study are presented in the next section.
%--------------------------Results on TMH data--------------------------
%------------------RFA 2 data
\begin{figure}[h!]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template1.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 1}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template2.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 2}
    \end{subfigure}
     \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template3.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 3}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template4.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 4}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template5.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 5}
    \end{subfigure}
              \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template6_color.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 6}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template7.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 7}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/colorTestIm_ablation.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 8}
    \end{subfigure}     
     \caption{Radio-frequency ablation dataset: One of the slices (512$\times$512) from each of the 8 scan volumes of a longitudinal study dataset of the liver. Note that in volumes (a) through (g), the needle (shown in red in (f)) approaches the target tumor. (h) the organ after the ablation: this slice is displayed on a separate intensity scale to enable proper viewing of the region marked in green that shows the after-effects of ablating the tumor.} 
\label{fig:RFA2_test_object-prior}
\end{figure}

@@

\subsubsection{\textbf{Okra dataset}}
\label{Sec:okra_spatially_varying}
One of the slices from the reconstructed volumes 
is shown in Figure.~\ref{fig:okra_3D_results}. The red and green 3D
RoI in the video and images show the regions where new changes are
present. The zoomed-in images around the major region of change (red
RoI) is shown in Figure~\ref{fig:okra_zoomed_3D_results}. As in the
test, the reconstruction by spatially-varying method shows the absence
of the deformity and better removal of sub-sampling artefacts when
compared to FDK and l1-ls. This is also seen in the SSIM values in
Table~\ref{table:okra_ssim}.

@@

Earlier in Sec.~\ref{sec:tmh}, we discussed that the imaging protocols fell under two categories depending on the final goal (tracking or observing details): very few-view imaging and moderate view imaging. \textbf{Ideally, we may prefer to gradually increase from few-view to moderate views as the probe gradually approaches the tumor site.} For the reconstruction of $n^{\textrm{th}}$ slice i.e., slice imaged at time $t=n$, the few-view reconstructions of the previously acquired slices can be used as object-prior. However, the first scan must always be taken with large number of views because it acts as a reference template to start with. Table.~\ref{tab:realistic_imaging_protocol} summarizes this protocol for the dataset of Fig~\ref{fig:RFA2_test_object-prior}, and Fig.~\ref{fig:RFA2_new_protocol} shows the reconstructions when this realistic protocol is used.

>>>>

 When the number of views is significantly reduced (this paper), the
 sparsity prior alone is not sufficient.  In such cases additional
 information (called the `object-prior') specific to the current
 object being scanned (called the `test' henceforth) is useful in
 further improving the reconstruction.  While this has been done in
 the context of dynamic CT-scans (as discussed in
 Section~\ref{sec:related}), in this paper we use a set $S$ of
 previous scans of the same object. This is relevant in a longitudinal
 context: the acquisition of sequential CT scans of the same subject
 in order to track time-evolving changes within the subject's
 interior. As seen in Fig.~\ref{fig:story} (last column, 30 views) our
 method uses the prior scans \emph{uniformly} by creating an
 eigenspace representation (Section~\ref{sec:method_uniform_prior}).

 However, with the use of such an object prior, a new challenge
 emerges. The prior set $S$ may potentially overwhelm the necessary
 details, and when several prior scans are available, finding the
 right prior is an issue. We therefore seek an algorithm that
 estimates the location and magnitude of new changes in the
 (unknown) test. As we show in
 Section~\ref{sec:method_spatially_varying_prior}, this eventually
 prevents the prior from adversely affecting the reconstruction of new
 regions in the test. We refer to this method as a
 \textit{spatially-varying} prior-based reconstruction routine, that
 still uses \textit{all} the previous scans in the set $S$, without
 using simply the prior scan, or endeavoring to find the ``right''
 prior.

@@

In this section, we survey the existing literature on tomographic reconstruction from undersampled measurements using object priors. In most cases, the object prior consists of an object reconstructed to high quality from a large number of projection angles. One of the earliest pieces of work in this area is the well-cited PICCS method~\cite{PICCS}, which enforces a robust-norm based similarity between the test and the object prior, in addition to a sparsity prior on the test. However in such a technique, there is no inherent mechanism to prevent the prior from overwhelming the current reconstruction, and choice of relative weights between the object and sparsity priors is difficult. A closely related method called PIRPLE was proposed in~\cite{pirple}, with additional steps to register the object prior and the test and a somewhat more flexible combination of data fidelity terms, sparsity prior and object prior. However, this method too has similar limitations as PICCS. In some pieces of work, the object prior is projected in the same angles as the current set of test measurements, and a set of projection differences is computed. These projection differences are essentially tomographic projections of the difference between the unknown test and the known object prior. Hence, a variety of reconstruction algorithms can be used to estimate such a difference image which can then be added to the object prior to yield the final estimate of the test. Such methods have been proposed in~\cite{Pourmorteza2015,Lee2012}. However, the reconstruction of the difference image will inherently contain artifacts due to a sparse set of projections, and these artifacts will appear in the final reconstruction as there is no mechanism to mitigate them (unlike the method that we propose). This can be seen in Sec.7 of our supplementary material~\cite{supp_paper}, where we have compared our method with~\cite{Lee2012}. In some recent techniques such as~\cite{Marjolein2016}, additional knowledge regarding regions of most likely change between the test and object prior is assumed, and lower weights are assigned to the prior in those regions. However, this is not done in a data driven manner and relies on additional expert domain knowledge which may not be available or which may be infeasible to acquire. There exists a fairly large-sized body of literature for reconstruction of structurally changing objects. In such cases, it may be infeasible to gather a large number of projections at every time instant. Moreover, there is additional temporal redundancy available which can reduce the required sample complexity for good reconstruction, and at the same time, tracking structural changes across time is of paramount importance. Many interesting approaches for this have been proposed in~\cite{daniil2015,koen2020, Van2015,HaoGao,Van2014}. These approaches variously use 4D sparsifying transforms for exploiting spatio-temporal redundancy~\cite{daniil2015}, robust principal components analysis assuming that the object consists of a large static part (background) and a sparse moving part (foreground)~\cite{HaoGao}, optical flow computations between projections (at different time instances) of the dynamic nonrigid object~\cite{koen2020} with additional key-frame and residual modelling in~\cite{Guangming2019}, spline-based models for tracking the regions of change in an SIRT framework~\cite{Van2014}, or knowledge of fluid models (piecewise constant shapes of attenuation curves in the dynamic region) in an SIRT framework~\cite{HaoGao}. However, all these techniques~\cite{daniil2015,koen2020,Van2015,HaoGao,Van2014} essentially reconstruct all the different stages of the object simultaneously, and use the extra redundancy across time simultaneously. In the scenario of a longitudinal study (where projection measurements are acquired at instants that are several weeks apart) which we consider in this paper, such approaches are not feasible. A longitudinal study will necessarily require reconstruction results after acquisition of each set of measurements. 
In this paper, we present a technique that makes use of multiple object priors, corresponding to high quality reconstructions of similar objects across time. We use these priors to improve the reconstruction of the test from largely undersampled measurements. In particular, we define the regions of change between the test and the eigenspace spanned by the previous high quality object priors in a purely data driven manner. Very importantly, our technique clearly differentiates between genuine structural changes and ``changes'' that appear due to undersampling artefacts.  Besides this, since we use a statistical model with multiple object priors, we avoid the problem of selecting an appropriate single object prior unlike~\cite{PICCS,Pourmorteza2015,Lee2012,pirple,Marjolein2016} and also make use of the additional information available in the multiple priors.

@@

 \begin{figure}[h]
\centering
	\includegraphics[width=0.5\textwidth]{../images/prior_cmb_edited.png}
        \caption{Overview of our work. The choice of the number of
          measurement views and the type of reconstruction is driven
          by the goal in the application under consideration.  (a) When
          our goal is relatively simple, such as tracking the location
          of new changes while simultaneously reducing sub-sampling
          artefacts, we propose acquiring measurements from an
          extremely small number of views (`few-view' imaging) and
          using uniform prior based reconstruction. (b) When our goal
          becomes more ambitious, such as observing details of the new
          changes while simultaneously reducing sub-sampling
          artefacts, we propose acquiring measurements from a
          slightly higher number of views (`moderate-view' imaging)
          and using spatially-varying prior-based reconstruction. In
          either case, the number of views is lower than what is
          conventionally used.}
 \label{fig:prior_overview}
\end{figure}

In summary, the key contributions are
\begin{itemize}
\item We create new 3D biological datasets (Section~\ref{sec:datasets}) and
  present results on real cone-beam projections. Our datasets and code
  will be made available to the community.
  \item As seen in Fig.~\ref{fig:story}, we demonstrate the use of
    uniform priors when very few views are used. More results appear
    in Section~\ref{sec:results_uniform_prior} and in the
    supplementary material.
  \item We discuss the scenarios where a uniform object prior is
    sufficient for the application at hand, and the scenarios for
    which it will fail. This leads to the idea of a weights map
    designed to depict the location and strength of new changes at
    every voxel. A novel algorithm is presented to build this map from
    sub-sampled measurements of the test and a set of high quality
    templates. Once the weights map is built, it is used for accurate
    reconstruction of those changes in the test that are absent in all
    of the templates. Results appear in
    Section~\ref{sec:results_spatially_varying_prior}.
  \item We show the efficacy of our results in a real-life medical
    longitudinal studies with data obtained in a clinical setting from
    a live teaching and
    research hospital 
\end{itemize}

@@

 \textbf{Okra:} This dataset is that of an Okra specimen consisting of its five
scans (Fig.~\ref{fig:object-prior_test_okra}). The measurements
consisted of real circular cone-beam projections from 450 views, each of size
336$\times$156. The corresponding size of the reconstructed volume is
338$\times$338$\times$123. Prior to the first scan, two holes were
drilled on the surface of the specimen. This was followed by four
scans, each after introducing one new cut. The specimen was kept in
the same position throughout the acquisitions. In cases where such an
alignment is not present, all the template volumes must be pre-aligned
before computing the eigenspace. The test must be registered to the
object-prior after its preliminary pilot reconstruction. The ground
truth consists of volumes reconstructed using the Feldkamp-Davis-Kress
(FDK) algorithm~\cite{FDK} from the full set of 450 view
projections. The test volume was reconstructed from a partial set of
45 projections, i.e., $10\%$ of the projection views from which ground
truth was reconstructed.\\
 %--------------------------------okra data-------------------------------
\begin{figure}[!h]
    \begin{subfigure}[b]{0.18\linewidth}
        \includegraphics[width=\textwidth]{../images/okra/template1_marked.png}
\captionsetup{labelformat=empty}       
 \caption{}
    \end{subfigure}
    \begin{subfigure}[b]{0.18\linewidth}
        \includegraphics[width=\textwidth]{../images/okra/template2_marked.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.18\linewidth}
        \includegraphics[width=\textwidth]{../images/okra/template3_marked.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.18\linewidth}
        \includegraphics[width=\textwidth]{../images/okra/template4_marked.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.176\linewidth}
        \fcolorbox{blue}{blue}{\includegraphics[width=\textwidth]{../images/okra/test_marked.png}}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
     \caption{Okra 3D dataset: One slice each from the previously scanned objects (the
       first four from the left), and one from the test volume
       (extreme right). In the regions marked in red and green, while
       all slices have deformities, the test
       has none.}
\label{fig:object-prior_test_okra}
\end{figure}
%----------------------------------------------------

\textbf{Potato:} This dataset consisted of four scans of the humble potato (Fig.~\ref{fig:object-prior_test_potato_A}). Measurements from each scan consisted of real circular cone-beam projections from 900 views, each of size 150$\times$150. The corresponding size of the reconstructed volume is 150$\times$150$\times$100. While the first scan was taken of the undistorted potato, subsequent scans were taken of the same specimen, each time after drilling a new hole halfway into the potato.  Projections were obtained using circular cone beam geometry.  %Further, this alignment can be refined following every update of the reconstructed test volume.
The ground truth consists of FDK reconstructions from the full set of acquired measurements from 900 projection views. The test volume was reconstructed using measurements from 45 projection views, i.e., $5\%$ of the projection views from which ground truth was reconstructed.
%---------------------------------------
\begin{figure}[!h]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/template_1.png}
\captionsetup{labelformat=empty}       
 \caption{}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/template_2.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/template_3.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.235\linewidth}
        \fcolorbox{blue}{blue}{\includegraphics[width=\textwidth]{../images/potato/testIm_color.png}}
\captionsetup{labelformat=empty}
\caption{}
\label{fig:potato3D_test}
     \end{subfigure}
      \caption{Potato 3D dataset: One slice each from the previously scanned objects (the first three from left) and a slice from the test 
        volume (extreme right). Notice the appearance of the fourth
        hole in the test slice. }
\label{fig:object-prior_test_potato_A}
\end{figure}

\subsection{Simulated measurements from real-life reconstructed volumes}
\textbf{Sprouts:} This 3D dataset, which was also obtained at ANU, consists of six reconstructed
volumes corresponding to six scans of an in-vivo sprout specimen
imaged at its various stages of growth
(Fig.~\ref{fig:object-prior_test_sprouts}). Projections were generated
from the given volume of size 130$\times$130$\times$130. The ground
truth consists of FDK reconstructed volumes from a set of 1800 view
projections. The test volume was reconstructed from partial set of 45
projections, i.e., $2.5\%$ of the projection views from which ground
truth was reconstructed.\\
%-------------------------sprouts data

\begin{figure}[!h]
    \begin{subfigure}[b]{0.3\linewidth}
        \includegraphics[width=\textwidth]{../images/sprouts/template_1.png}
\captionsetup{labelformat=empty}
        \caption{}
    \end{subfigure}
\quad
    \begin{subfigure}[b]{0.3\linewidth}
        \includegraphics[width=\textwidth]{../images/sprouts/template_2.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
\quad
    \begin{subfigure}[b]{0.3\linewidth}
        \includegraphics[width=\textwidth]{../images/sprouts/template_3.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}\\
    \begin{subfigure}[b]{0.3\linewidth}
        \includegraphics[width=\textwidth]{../images/sprouts/template_4.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
\quad
    \begin{subfigure}[b]{0.3\linewidth}
        \includegraphics[width=\textwidth]{../images/sprouts/template_5.png}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
\quad
    \begin{subfigure}[b]{0.29\linewidth}
        \fcolorbox{blue}{blue}{\includegraphics[width=\textwidth]{../images/sprouts/testIm_red.png}}
\captionsetup{labelformat=empty}
        \caption{}
     \end{subfigure}
      \caption{Sprouts 3D dataset: One slice each from the previously scanned objects (the first five from left) and a slice from the test (extreme
        right). Notice the structure within the red box in the test that is different from all of the other images.}
\label{fig:object-prior_test_sprouts}
\addtolength{\textfloatsep}{-0.8cm}
\end{figure}

\textbf{Radio-frequency ablation:} We have used this dataset (Fig.~\ref{fig:RFA2_test_object-prior} to
illustrate an application where \textit{both} uniform and spatially-varying priors are
useful. The data~\footnote{Source: Tata Memorial
  Centre~\cite{tmh}, Parel, Mumbai.  This is the national
  comprehensive centre for the prevention, treatment, education and
  research in cancer, and is recognized as one of the leading cancer
  centres in India.} consists of simulated parallel beam measurements of 2D slices
of the organ (liver) at different stages of the scan. The details of this study are presented in the next section.
%--------------------------Results on TMH data--------------------------
%------------------RFA 2 data
\begin{figure}[h!]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template1.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 1}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template2.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 2}
    \end{subfigure}
     \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template3.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 3}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template4.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 4}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template5.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 5}
    \end{subfigure}
              \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template6_color.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 6}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template7.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 7}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/colorTestIm_ablation.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 8}
    \end{subfigure}     
     \caption{Radio-frequency ablation dataset: One of the slices (512$\times$512) from each of the 8 scan volumes of a longitudinal study dataset of the liver. Note that in volumes (a) through (g), the needle (shown in red in (f)) approaches the target tumor. (h) the organ after the ablation: this slice is displayed on a separate intensity scale to enable proper viewing of the region marked in green that shows the after-effects of ablating the tumor.} 
\label{fig:RFA2_test_object-prior}
\end{figure}

@@

\subsubsection{\textbf{Okra dataset}}
\label{Sec:okra_spatially_varying}
One of the slices from the reconstructed volumes 
is shown in Fig.~\ref{fig:okra_3D_results}. The red and green 3D
RoI in the video and images show the regions where new changes are
present. The zoomed-in images around the major region of change (red
RoI) is shown in Fig.~\ref{fig:okra_zoomed_3D_results}. As in the
test, the reconstruction by spatially-varying method shows the absence
of the deformity and better removal of sub-sampling artefacts when
compared to FDK and l1-ls. This is also seen in the SSIM values in
Table~\ref{table:okra_ssim}.

@@
